Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-23 paper/icml_drau/advup.tex
.

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
\TODO{
For the algorithm to admit a continuous-time limit, the discrete-time trajectories
of parameters must converge to well-defined trajectories as $\deltat$ goes to
$0$.
% The optimization steps should not be too large,
% in which case parameter trajectories could jump in a single gradient step,
% or too small, in which case the trajectories would become stationary as
% $\deltat$ goes to $0$.
This in turn imposes precise conditions on the scalings of the
learning rates.}% of algorithms robust to changes of the time discretization.  }

>>>>

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the algorithm to admit a continuous-time limit, the discrete-time trajectories
of parameters must converge to well-defined trajectories as $\deltat$ goes to
$0$.
% The optimization steps should not be too large,
% in which case parameter trajectories could jump in a single gradient step,
% or too small, in which case the trajectories would become stationary as
% $\deltat$ goes to $0$.
This in turn imposes precise conditions on the scalings of the
learning rates.% of algorithms robust to changes of the time discretization.  }

### 2019-01-23 paper/icml_drau/experiments.tex
.

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Algo 1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. We tested two different setups for DDPG and DQN.
In one, we scaled the discount factor (to avoid shortsightedness with small $\deltat$), but
left all other hyperparameters constant across time discretizations.
In the other, we used the properly rescaled discount
factor and reward from Eq.~\eqref{eq:def-gamma},
as well as $O(\deltat)$ learning rates for RMSProp.  The former yields slightly
better results, and is presented in the main text. Experiments on the second
variant are presented in the
supplementary material.

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
return results are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
normalized epochs (physical time), i.e. the true number of epochs multiplied by the
discretization timestep.

@@

We plot the learnt policy at several instants in physical time (or
normalized epochs) during training, for various time discretizations
$\deltat$, for both DAU and DDPG. With DAU, the agent's policy quickly
converges for every time discretization without specific hyperparameter
tuning. On the contrary, with DDPG, the agent is only able to learn for
the two largest time discretizations; DDPG learning gets harder as $\deltat$ decreases.

\paragraph{Quantitative experiments.}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum,
Cartpole, BipedalWalker, Ant, and Half-Cheetah environments from OpenAI
Gym. For Pendulum and Ant, DDPG is able to learn for the highest
$\deltat$ but fails for the others, while DAU remains reasonably
invariant to the time discretization. On BipedalWalker, DDPG is never
able to learn while DAU is achieving decent performance for all the
$\deltat$'s. On Cartpole and Half-Cheetah, the results are less clear. On
Cartpole, noise dominates, which makes interpretation difficult. On Half-Cheetah, DAU is not clearly more invariant to time discretization than
DDPG. This could be explained by the multiple suboptimal regimes that
coexist in the Half-Cheetah environment (walking on the head, walking on
the back).
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/cartpole_lc.png}
%   \caption{Cartpole}
%   \label{fig:cartpole-lc}
% \end{figure}

>>>>

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Alg.~1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. We tested two different setups for DDPG and DQN.
In one, we scaled the discount factor (to avoid shortsightedness with small $\deltat$), but
left all other hyperparameters constant across time discretizations.
In the other, we used the properly rescaled discount
factor and reward from Eq.~\eqref{eq:def-gamma},
as well as $O(\deltat)$ learning rates for RMSProp.  The former yields slightly
better results, and is presented in the main text. Experiments on the second
variant are presented in the
supplementary material.

Let us stress that the quantities plotted are rescaled to make comparison
possible across different timesteps. For example,
return results are given in term of discretized return $R_\deltat$ as defined in \eqref{eq:discretized-return},\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment-specific
details are given in the supplementary material.} and, most notably, time elapsed is always given in
physical time, i.e. the amount of time that the agent spent interacting with the environment (this is not the computational time).

@@

We plot the learnt policy at several instants in physical time (or
normalized epochs) during training, for various time discretizations
$\deltat$, for both DAU and DDPG. With DAU, the agent's policy and value function quickly
converge for every time discretization without specific
tuning. On the contrary, with DDPG, learning of both value function
and policy greatly vary from one discretization to another.

\paragraph{Quantitative experiments.}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum,
Cartpole, BipedalWalker, Ant, and Half-Cheetah environments from OpenAI Gym. On
Pendulum, Bipedal Walker and Ant, DAU is quite robust to variations of
$\deltat$ and displays reasonable performances in all cases. On the other hand,
DDPG's performance varies with $\deltat$; performance are either degraded as
$\deltat$ decreases (Ant, Cheetah), or become more variable as learning
progresses (Pendulum) with small $\deltat$'s. On Cartpole, noise dominates,
which makes interpretation difficult. On Half-Cheetah, DAU is not clearly
invariant to time discretization. This could be explained by the multiple
suboptimal regimes that coexist in the Half-Cheetah environment (walking on the
head, walking on the back).
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/cartpole_lc.png}
%   \caption{Cartpole}
%   \label{fig:cartpole-lc}
% \end{figure}
